Inattentive responding can induce spurious associations between task behavior and symptom measures

### 2023-04-04 manuscripts/main.tex
checkpointing manuscript progress

% load packages
\usepackage[utf8]{inputenc}
\usepackage[a4paper, margin=0.8in]{geometry}
\usepackage{authblk}
\usepackage[english]{babel}
\usepackage{graphicx}
\usepackage{indentfirst}
\usepackage{xr-hyper} % for cross-referencing between documents
\usepackage{hyperref}
\usepackage{float}
\usepackage{amsmath}
\usepackage[backend=biber,style=nature]{biblatex}
\usepackage[flushmargin]{footmisc}
\usepackage[flushleft]{threeparttable}
\usepackage{array}
\usepackage[font=small, labelfont=it]{caption}
\usepackage{booktabs}
\usepackage{xcolor,soul}

% specify footnote margins
\renewcommand\footnotelayout{%
  \advance\leftskip 0.7cm
  \advance\rightskip 0.7cm
 }

% setup references
\defbibheading{main}{}
\addbibresource[label=main]{main.bib}
\renewcommand*{\bibfont}{\small}

@@

@@

\begin{document}
%TC:ignore

@@

% make abstract
\abstract{
A common research design in the field of computational psychiatry involves leveraging the power of online participant recruitment to assess correlations between behavior in cognitive tasks and the self-reported severity of psychiatric symptoms in large, diverse samples. Although large online samples have many advantages for psychiatric research, some potential pitfalls of this research design are not widely understood. Here we detail circumstances in which entirely spurious correlations may arise between task behavior and symptom severity as a result of inadequate screening of careless or low-effort responding on psychiatric symptom surveys. Specifically, since many psychiatric symptom surveys have asymmetric ground-truth score distributions in the general population, participants who respond carelessly on these surveys will show apparently elevated symptom levels. If these participants are similarly careless in their task performance, and are not excluded from analysis, this may result in a spurious association between greater symptom scores and worse behavioral task performance. Here, we demonstrate exactly this pattern of results in two independent samples of participants (total $N = 779$) recruited online to complete a self-report symptom battery and one of two common cognitive tasks. We show that many behavior-symptom correlations are entirely abolished when participants flagged for careless responding on surveys are excluded from analysis. We also show that exclusion based on task performance alone is often insufficient to prevent these spurious correlations. Of note, we demonstrate that false-positive rates for these spurious correlations \emph{increase} with sample size, contrary to common assumptions. We offer guidance on how researchers using this general experimental design can guard against this issue in future research; in particular, we recommend the adoption of screening methods for self-report measures that are currently uncommon in this field.
}

@@

%TC:endignore

\begin{refsection}[main]
\section*{Introduction}

@@

\subsection*{Screening for C/IE responding is common for task behavior but not for self-report surveys}

@@

Among studies that we reviewed, approximately 80\% (39/49) used at least one method to identify C/IE responding in task behavior (see Table \ref{tab:screening}). Of these, just over half relied on a single screening method, with considerable heterogeneity in behavior screening methods across studies. Most common (46\% of all studies) was identifying participants whose performance was statistically indistinguishable from chance-level on some measure of accuracy. Almost as common (38\%) was screening based on low response variability (i.e., excluding participants who predominantly responded in the same fashion across trials, such as using only a single response key).

@@

\subsection*{Careless participants appear symptomatic when the overall level of symptom endorsement is low}

@@

We hypothesise that spurious behavior-symptom correlations may emerge due to a mean-shift in the average level of symptom endorsement in participants engaging in C/IE responding relative to attentive participants. In turn, a mean-shift is expected to occur when the overall rate of symptom endorsement is low; that is, comparably higher scores are more likely for C/IE participants responding at random on a right skewed questionnaire. In line with our predictions, the average level of symptom endorsement was noticeably exaggerated in C/IE-responding participants for the symptom measures where symptom scores were most positively-skewed (7-up, 7-down, GAD-7; see Figure \ref{fig:distributions}). In contrast, where there was higher rates of symptom endorsement overall, the distributions of symptom scores between the two groups of participants were less noticeably distinct. Permutation testing confirmed that observed mean-shifts in symptom scores for C/IE participants were statistically significant for the majority of symptom measures (Table \ref{tab:validation}).

Hereafter, we use the infrequency-item method as a primary means of identifying C/IE responding in our data. To verify this approach, we conducted three validation analyses. The first analysis compared estimated internal consistency of self-report measures between the C/IE and attentive groups. The logic is that, if C/IE responding manifests as a tendency to respond randomly, we should expect to see a decrease in the consistency of a measure in the C/IE responding group \cite{maniaci2014caring, desimone2018dirty, desimone2018differential}. In line with this reasoning, we observed a reduction in Cronbach's $\alpha$ in the C/IE group for the majority of survey instruments (Table \ref{tab:validation}). A permutation test confirmed that the average decrease in internal consistency across measures was greater than would be expected by chance given the difference in participant numbers between groups (two-tailed, paired-samples $t$-test: $t(6)=3.689$, $p=0.021$, $d=1.506$, 95\% CI = [0.048, 0.141]).

@@

\subsection*{Spurious symptom-behavior correlations produced by C/IE responding}

@@

\subsection*{Individuals with and without diagnosed psychiatric disorders fail attention checks at similar rates}

>>>>

% load packages
\usepackage[utf8]{inputenc}
\usepackage[a4paper, margin=0.8in]{geometry}
\usepackage{authblk}
\usepackage[english]{babel}
\usepackage{graphicx}
\usepackage{indentfirst}
\usepackage{hyperref}
\usepackage{float}
\usepackage{amsmath}
\usepackage[backend=biber,style=nature]{biblatex}
\usepackage[flushmargin]{footmisc}
\usepackage[flushleft]{threeparttable}
\usepackage{array}
\usepackage[font=small, labelfont=it]{caption}
\usepackage{booktabs}
\usepackage{xcolor,soul}

@@

% ----------------------------------%
% Define references
% ----------------------------------%

% setup references
\renewcommand*{\bibfont}{\small}

% define references
\begin{filecontents}{main.bib}
@article{stewart2017crowdsourcing,
  title={Crowdsourcing samples in cognitive science},
  author={Stewart, Neil and Chandler, Jesse and Paolacci, Gabriele},
  journal={Trends in cognitive sciences},
  volume={21},
  number={10},
  pages={736--748},
  year={2017},
  publisher={Elsevier}
}

@article{chandler2016conducting,
  title={Conducting clinical research using crowdsourced convenience samples},
  author={Chandler, Jesse and Shapiro, Danielle},
  journal={Annual review of clinical psychology},
  volume={12},
  year={2016}
}

@article{gillan2016taking,
  title={Taking psychiatry research online},
  author={Gillan, Claire M and Daw, Nathaniel D},
  journal={Neuron},
  volume={91},
  number={1},
  pages={19--23},
  year={2016},
  publisher={Elsevier}
}

@article{rutledge2019machine,
  title={Machine learning and big data in psychiatry: toward clinical applications},
  author={Rutledge, Robb B and Chekroud, Adam M and Huys, Quentin JM},
  journal={Current opinion in neurobiology},
  volume={55},
  pages={152--159},
  year={2019},
  publisher={Elsevier}
}

@article{strickland2019use,
  title={The use of crowdsourcing in addiction science research: Amazon Mechanical Turk.},
  author={Strickland, Justin C and Stoops, William W},
  journal={Experimental and Clinical Psychopharmacology},
  volume={27},
  number={1},
  pages={1},
  year={2019},
  publisher={American Psychological Association}
}

@article{enkavi2019large,
  title={Large-scale analysis of test--retest reliabilities of self-regulation measures},
  author={Enkavi, A Zeynep and Eisenberg, Ian W and Bissett, Patrick G and Mazza, Gina L and MacKinnon, David P and Marsch, Lisa A and Poldrack, Russell A},
  journal={Proceedings of the National Academy of Sciences},
  volume={116},
  number={12},
  pages={5472--5477},
  year={2019},
  publisher={National Acad Sciences}
}

@article{kothe2019retention,
  title={Retention of participants recruited to a one-year longitudinal study via Prolific},
  author={Kothe, Emily and Ling, Mathew},
  year={2019},
  journal={PsyArXiv},
  publisher={PsyArXiv}
}

@article{chandler2020participant,
  title={Participant carelessness and fraud: Consequences for clinical research and potential solutions.},
  author={Chandler, Jesse and Sisso, Itay and Shapiro, Danielle},
  journal={Journal of Abnormal Psychology},
  volume={129},
  number={1},
  pages={49},
  year={2020},
  publisher={American Psychological Association}
}

@article{ophir2020turker,
  title={The turker blues: Hidden factors behind increased depression rates among Amazon’s Mechanical Turkers},
  author={Ophir, Yaakov and Sisso, Itay and Asterhan, Christa SC and Tikochinski, Refael and Reichart, Roi},
  journal={Clinical Psychological Science},
  volume={8},
  number={1},
  pages={65--83},
  year={2020},
  publisher={Sage Publications Sage CA: Los Angeles, CA}
}

@article{curran2016methods,
  title={Methods for the detection of carelessly invalid responses in survey data},
  author={Curran, Paul G},
  journal={Journal of Experimental Social Psychology},
  volume={66},
  pages={4--19},
  year={2016},
  publisher={Elsevier}
}

@article{desimone2018dirty,
  title={Dirty data: The effects of screening respondents who provide low-quality data in survey research},
  author={DeSimone, Justin A and Harms, PD},
  journal={Journal of Business and Psychology},
  volume={33},
  number={5},
  pages={559--577},
  year={2018},
  publisher={Springer}
}

@article{desimone2018differential,
  title={The differential impacts of two forms of insufficient effort responding},
  author={DeSimone, Justin A and DeSimone, Alice J and Harms, PD and Wood, Dustin},
  journal={Applied Psychology},
  volume={67},
  number={2},
  pages={309--338},
  year={2018},
  publisher={Wiley Online Library}
}

@article{hauser2016attentive,
  title={Attentive Turkers: MTurk participants perform better on online attention checks than do subject pool participants},
  author={Hauser, David J and Schwarz, Norbert},
  journal={Behavior research methods},
  volume={48},
  number={1},
  pages={400--407},
  year={2016},
  publisher={Springer}
}

@article{thomas2017validity,
  title={Validity and Mechanical Turk: An assessment of exclusion methods and interactive experiments},
  author={Thomas, Kyle A and Clifford, Scott},
  journal={Computers in Human Behavior},
  volume={77},
  pages={184--197},
  year={2017},
  publisher={Elsevier}
}

@article{barends2019noncompliant,
  title={Noncompliant responding: Comparing exclusion criteria in MTurk personality research to improve data quality},
  author={Barends, Ard J and de Vries, Reinout E},
  journal={Personality and individual differences},
  volume={143},
  pages={84--89},
  year={2019},
  publisher={Elsevier}
}

@article{mukherjee_reward_2020,
  title={Reward and punishment reversal-learning in major depressive disorder},
  author={Mukherjee, D and Filipwicz, A L S and Vo, K and Satterthwaite, T D and Kable, J W},
  journal={Journal of Abnormal Psychology},
  volume={129},
  number={8},
  pages={810--823},
  year={2020}
}

@article{waltz2007probabilistic,
  title={Probabilistic reversal learning impairments in schizophrenia: further evidence of orbitofrontal dysfunction},
  author={Waltz, James A and Gold, James M},
  journal={Schizophrenia Research},
  volume={93},
  number={1-3},
  pages={296--303},
  year={2007},
  publisher={Elsevier}
}

@article{lowe2008validation,
  title={Validation and standardization of the Generalized Anxiety Disorder Screener (GAD-7) in the general population},
  author={Lowe, Bernd and Decker, Oliver and Muller, Stefanie and Brahler, Elmar and Schellberg, Dieter and Herzog, Wolfgang and Herzberg, Philipp Yorck},
  journal={Medical care},
  pages={266--274},
  year={2008},
  publisher={JSTOR}
}

@article{hinz2017psychometric,
  title={Psychometric evaluation of the Generalized Anxiety Disorder Screener GAD-7, based on a large German general population sample},
  author={Hinz, Andreas and Klein, Annette M and Brahler, Elmar and Glaesmer, Heide and Luck, Tobias and Riedel-Heller, Steffi G and Wirkner, Kerstin and Hilbert, Anja},
  journal={Journal of affective disorders},
  volume={210},
  pages={338--344},
  year={2017},
  publisher={Elsevier}
}

@article{maydeu2006random,
  title={Random intercept item factor analysis.},
  author={Maydeu-Olivares, Albert and Coffman, Donna L},
  journal={Psychological methods},
  volume={11},
  number={4},
  pages={344},
  year={2006},
  publisher={American Psychological Association}
}

@article{deleeuw_2015_jspsych,
  title={{jsPsych: A JavaScript} library for creating behavioral experiments in a web browser. },
  author={de Leeuw, J R},
  journal={Behavior Research Methods},
  volume={47},
  number={1},
  pages={1--12},
  year={2015}
}

@article{youngstrom_2013_susd,
  title={The {7 Up 7 Down Inventory}: A 14-item measure of manic and depressive tendencies carved from the {General Behavior Inventory}},
  author={Youngstrom, Eric A and Murray, Greg and Johnson, Sheri L and Findling, Robert L},
  journal={Psychological Assessment},
  volume={25},
  number={4},
  pages={1377--1383},
  year={2013},
  publisher={American Psychological Association}
}

@article{depue1981behavioral,
  title={A behavioral paradigm for identifying persons at risk for bipolar depressive disorder: a conceptual framework and five validation studies},
  author={Depue, Richard A and Slater, Judith F and Wolfstetter-Kausch, Heidi and Klein, Daniel and Goplerud, Eric and Farr, David},
  journal={Journal of Abnormal Psychology},
  volume={90},
  number={5},
  pages={381--437},
  year={1981},
  publisher={American Psychological Association}
}

@article{spitzer2006brief,
  title={A brief measure for assessing generalized anxiety disorder: the {GAD-7}},
  author={Spitzer, Robert L and Kroenke, Kurt and Williams, Janet BW and Lowe, Bernd},
  journal={Archives of Internal Medicine},
  volume={166},
  number={10},
  pages={1092--1097},
  year={2006},
  publisher={American Medical Association}
}

@article{carver1994behavioral,
  title={Behavioral inhibition, behavioral activation, and affective responses to impending reward and punishment: the {BIS/BAS} scales.},
  author={Carver, Charles S and White, Teri L},
  journal={Journal of Personality and Social Psychology},
  volume={67},
  number={2},
  pages={319--333},
  year={1994},
  publisher={American Psychological Association}
}

@article{pagliaccio2016revising,
  title={Revising the {BIS/BAS} Scale to study development: measurement invariance and normative effects of age and sex from childhood through adulthood},
  author={Pagliaccio, David and Luking, Katherine R and Anokhin, Andrey P and Gotlib, Ian H and Hayden, Elizabeth P and Olino, Thomas M and Peng, Chun-Zi and Hajcak, Greg and Barch, Deanna M},
  journal={Psychological ssessment},
  volume={28},
  number={4},
  pages={429--442},
  year={2016},
  publisher={American Psychological Association}
}

@article{snaith1995scale,
  title={A scale for the assessment of hedonic tone: the Snaith--Hamilton Pleasure Scale},
  author={Snaith, RP and Hamilton, M and Morley, S and Humayan, A and Hargreaves, D and Trigwell, P},
  journal={The British Journal of Psychiatry},
  volume={167},
  number={1},
  pages={99--103},
  year={1995},
  publisher={Cambridge University Press}
}

@article{meyer1990development,
  title={Development and validation of the {Penn State Worry Questionnaire}},
  author={Meyer, Thomas J and Miller, Mark L and Metzger, Richard L and Borkovec, Thomas D},
  journal={Behaviour Research and Therapy},
  volume={28},
  number={6},
  pages={487--495},
  year={1990},
  publisher={Elsevier}
}

@article{kertz2014psychometric,
  title={Psychometric properties of abbreviated and ultra-brief versions of the {Penn State Worry Questionnaire}},
  author={Kertz, Sarah J and Lee, Josephine and Bjorgvinsson, Throstur},
  journal={Psychological Assessment},
  volume={26},
  number={4},
  pages={1146--1154},
  year={2014},
  publisher={American Psychological Association}
}

@article{youngstrom2020evaluating,
  title={Evaluating and Validating GBI Mania and Depression Short Forms for Self-Report of Mood Symptoms},
  author={Youngstrom, Eric A and Perez Algorta, Guillermo and Youngstrom, Jennifer Kogos and Frazier, Thomas W and Findling, Robert L},
  journal={Journal of Clinical Child \& Adolescent Psychology},
  pages={1--17},
  year={2020},
  publisher={Taylor \& Francis}
}

@article{franken2007assessment,
  title={The assessment of anhedonia in clinical and non-clinical populations: further validation of the Snaith--Hamilton Pleasure Scale (SHAPS)},
  author={Franken, Ingmar HA and Rassin, Eric and Muris, Peter},
  journal={Journal of affective disorders},
  volume={99},
  number={1-3},
  pages={83--89},
  year={2007},
  publisher={Elsevier}
}

@article{marjanovic2015inter,
  title={The inter-item standard deviation (ISD): An index that discriminates between conscientious and random responders},
  author={Marjanovic, Zdravko and Holden, Ronald and Struthers, Ward and Cribbie, Robert and Greenglass, Esther},
  journal={Personality and Individual Differences},
  volume={84},
  pages={79--83},
  year={2015},
  publisher={Elsevier}
}

@article{ward2018applying,
  title={Applying social psychology to prevent careless responding during online surveys},
  author={Ward, MK and Meade, Adam W},
  journal={Applied Psychology},
  volume={67},
  number={2},
  pages={231--263},
  year={2018},
  publisher={Wiley Online Library}
}

@article{emons2009detection,
  title={Detection and diagnosis of person misfit from patterns of summed polytomous item scores},
  author={Emons, Wilco HM},
  journal={Applied Psychological Measurement},
  volume={33},
  number={8},
  pages={599--619},
  year={2009},
  publisher={Sage Publications Sage CA: Los Angeles, CA}
}

@article{robinson2014inaccurate,
  title={Inaccurate estimation of disparities due to mischievous responders: Several suggestions to assess conclusions},
  author={Robinson-Cimpian, Joseph P},
  journal={Educational Researcher},
  volume={43},
  number={4},
  pages={171--185},
  year={2014},
  publisher={Sage Publications Sage CA: Los Angeles, CA}
}

@article{arias2020little,
  title={A little garbage in, lots of garbage out: Assessing the impact of careless responding in personality survey data},
  author={Arias, Victor B and Garrido, LE and Jenaro, C and Martinez-Molina, A and Arias, B},
  journal={Behavior Research Methods},
  pages={1--17},
  year={2020},
  publisher={Springer}
}

@article{huang2012detecting,
  title={Detecting and deterring insufficient effort responding to surveys},
  author={Huang, Jason L and Curran, Paul G and Keeney, Jessica and Poposki, Elizabeth M and DeShon, Richard P},
  journal={Journal of Business and Psychology},
  volume={27},
  number={1},
  pages={99--114},
  year={2012},
  publisher={Springer}
}

@article{maniaci2014caring,
  title={Caring about carelessness: Participant inattention and its effects on research},
  author={Maniaci, Michael R and Rogge, Ronald D},
  journal={Journal of Research in Personality},
  volume={48},
  pages={61--83},
  year={2014},
  publisher={Elsevier}
}

@article{huang2015detecting,
  title={Detecting insufficient effort responding with an infrequency scale: Evaluating validity and participant reactions},
  author={Huang, Jason L and Bowling, Nathan A and Liu, Mengqiao and Li, Yuhui},
  journal={Journal of Business and Psychology},
  volume={30},
  number={2},
  pages={299--311},
  year={2015},
  publisher={Springer}
}

@article{huang2015insufficient,
  title={Insufficient effort responding: Examining an insidious confound in survey data.},
  author={Huang, Jason L and Liu, Mengqiao and Bowling, Nathan A},
  journal={Journal of Applied Psychology},
  volume={100},
  number={3},
  pages={828},
  year={2015},
  publisher={American Psychological Association}
}

@article{huang2017computational,
  title={Computational dysfunctions in anxiety: Failure to differentiate signal from noise},
  author={Huang, He and Thompson, Wesley and Paulus, Martin P},
  journal={Biological psychiatry},
  volume={82},
  number={6},
  pages={440--446},
  year={2017},
  publisher={Elsevier}
}

@article{brolsma2020challenging,
  title={Challenging the negative learning bias hypothesis of depression: reversal learning in a naturalistic psychiatric sample},
  author={Brolsma, Sophie CA and Vrijsen, Janna N and Vassena, Eliana and Kandroodi, Mojtaba Rostami and Bergman, M Annemiek and van Eijndhoven, Philip F and Collard, Rose M and den Ouden, Hanneke EM and Schene, Aart H and Cools, Roshan},
  journal={Psychological Medicine},
  pages={1--11},
  year={2020},
  publisher={Cambridge University Press}
}

@article{ritschel2017neural,
  title={Neural correlates of altered feedback learning in women recovered from anorexia nervosa},
  author={Ritschel, Franziska and Geisler, Daniel and King, Joseph A and Bernardoni, Fabio and Seidel, Maria and Boehm, Ilka and Vettermann, Richard and Biemann, Ronald and Roessner, Veit and Smolka, Michael N and others},
  journal={Scientific reports},
  volume={7},
  number={1},
  pages={1--10},
  year={2017},
  publisher={Nature Publishing Group}
}

@article{kool2018mental,
  title={Mental labour},
  author={Kool, Wouter and Botvinick, Matthew},
  journal={Nature human behaviour},
  volume={2},
  number={12},
  pages={899--908},
  year={2018},
  publisher={Nature Publishing Group}
}

@article{elwert2014endogenous,
  title={Endogenous selection bias: The problem of conditioning on a collider variable},
  author={Elwert, Felix and Winship, Christopher},
  journal={Annual review of sociology},
  volume={40},
  pages={31--53},
  year={2014},
  publisher={Annual Reviews}
}

@ARTICLE{Kim2018-ev,
  title    = "Detecting random responders with infrequency scales using an
              error-balancing threshold",
  author   = "Kim, Dale S and McCabe, Connor J and Yamasaki, Brianna L and
              Louie, Kristine A and King, Kevin M",
  journal  = "Behav. Res. Methods",
  volume   =  50,
  number   =  5,
  pages    = "1960--1970",
  month    =  oct,
  year     =  2018,
  keywords = "Careless responding; Data cleaning; Infrequency scales; Random
              responding",
  language = "en"
}

@article{culbreth2016negative,
  title={Negative symptoms are associated with an increased subjective cost of cognitive effort.},
  author={Culbreth, Adam and Westbrook, Andrew and Barch, Deanna},
  journal={Journal of abnormal psychology},
  volume={125},
  number={4},
  pages={528},
  year={2016},
  publisher={American Psychological Association}
}

@article{kane2016individual,
  title={Individual differences in the executive control of attention, memory, and thought, and their associations with schizotypy.},
  author={Kane, Michael J and Meier, Matt E and Smeekens, Bridget A and Gross, Georgina M and Chun, Charlotte A and Silvia, Paul J and Kwapil, Thomas R},
  journal={Journal of Experimental Psychology: General},
  volume={145},
  number={8},
  pages={1017},
  year={2016},
  publisher={American Psychological Association}
}

@article{robison2017neurotic,
  title={The neurotic wandering mind: An individual differences investigation of neuroticism, mind-wandering, and executive control},
  author={Robison, Matthew K and Gath, Katherine I and Unsworth, Nash},
  journal={The Quarterly Journal of Experimental Psychology},
  volume={70},
  number={4},
  pages={649--663},
  year={2017},
  publisher={Taylor \& Francis}
}

@article{cohen2001impairments,
  title={Impairments of attention and effort among patients with major affective disorders},
  author={Cohen, Ronald and Lohr, Ilan and Paul, Robert and Boland, Robert},
  journal={The Journal of neuropsychiatry and clinical neurosciences},
  volume={13},
  number={3},
  pages={385--395},
  year={2001},
  publisher={Am Neuropsych Assoc}
}

@article{barch2015mechanisms,
  title={Mechanisms underlying motivational deficits in psychopathology: similarities and differences in depression and schizophrenia},
  author={Barch, Deanna M and Pagliaccio, David and Luking, Katherine},
  journal={Behavioral neuroscience of motivation},
  pages={411--449},
  year={2015},
  publisher={Springer}
}

@article{buchanan2018methods,
  title={Methods to detect low quality data and its implication for psychological research},
  author={Buchanan, Erin M and Scofield, John E},
  journal={Behavior research methods},
  volume={50},
  number={6},
  pages={2586--2596},
  year={2018},
  publisher={Springer}
}

@article{hunter2019excessive,
  title={Excessive deliberation in social anxiety},
  author={Hunter, Lindsay E and Meer, Elana A and Gillan, Claire M and Hsu, Ming and Daw, Nathaniel D},
  journal={bioRxiv},
  pages={522433},
  year={2019},
  publisher={Cold Spring Harbor Laboratory}
}

@article{litman2017turkprime,
  title={TurkPrime. com: A versatile crowdsourcing data acquisition platform for the behavioral sciences},
  author={Litman, Leib and Robinson, Jonathan and Abberbock, Tzvi},
  journal={Behavior research methods},
  volume={49},
  number={2},
  pages={433--442},
  year={2017},
  publisher={Springer}
}

@article{robinson2019tapped,
  title={Tapped out or barely tapped? Recommendations for how to harness the vast and largely unused potential of the Mechanical Turk participant pool},
  author={Robinson, Jonathan and Rosenzweig, Cheskie and Moss, Aaron J and Litman, Leib},
  journal={PloS one},
  volume={14},
  number={12},
  pages={e0226394},
  year={2019},
  publisher={Public Library of Science San Francisco, CA USA}
}

@online{cloudresearch_2020,
    author = "Leib Litman",
    title = "New Solutions Dramatically Improve Research Data Quality on {MTurk}",
    url  = "https://www.cloudresearch.com/resources/blog/new-tools-improve-research-data-quality-mturk/",
    addendum = "(Accessed: 2021-02-23)",
    month = jul,
    year = 2020
}

@article{winkler2014permutation,
  title={Permutation inference for the general linear model},
  author={Winkler, Anderson M and Ridgway, Gerard R and Webster, Matthew A and Smith, Stephen M and Nichols, Thomas E},
  journal={Neuroimage},
  volume={92},
  pages={381--397},
  year={2014},
  publisher={Elsevier}
}

@article{grant2009typology,
  title={A typology of reviews: an analysis of 14 review types and associated methodologies},
  author={Grant, Maria J and Booth, Andrew},
  journal={Health information \& libraries journal},
  volume={26},
  number={2},
  pages={91--108},
  year={2009},
  publisher={Wiley Online Library}
}

@article{niv2012neural,
  title={Neural prediction errors reveal a risk-sensitive reinforcement-learning process in the human brain},
  author={Niv, Yael and Edlund, Jeffrey A and Dayan, Peter and O'Doherty, John P},
  journal={Journal of Neuroscience},
  volume={32},
  number={2},
  pages={551--562},
  year={2012},
  publisher={Soc Neuroscience}
}

@online{stan,
    title={Stan Modeling Language Users Guide and Reference Manual},
    author={{Stan Development Team}},
    year={2021},
     url  = "https://mc-stan.org",
}

@article{wilcox2018guide,
  title={A guide to robust statistical methods in neuroscience},
  author={Wilcox, Rand R and Rousselet, Guillaume A},
  journal={Current protocols in neuroscience},
  volume={82},
  number={1},
  pages={8--42},
  year={2018},
  publisher={Wiley Online Library}
}

@article{tomitaka2018distributional,
  title={Distributional patterns of item responses and total scores on the PHQ-9 in the general population: data from the National Health and Nutrition Examination Survey},
  author={Tomitaka, Shinichiro and Kawasaki, Yohei and Ide, Kazuki and Akutagawa, Maiko and Yamada, Hiroshi and Ono, Yutaka and Furukawa, Toshiaki A},
  journal={BMC psychiatry},
  volume={18},
  number={1},
  pages={1--9},
  year={2018},
  publisher={Springer}
}

@article{eldar2015interaction,
  title={Interaction between emotional state and learning underlies mood instability},
  author={Eldar, Eran and Niv, Yael},
  journal={Nature communications},
  volume={6},
  number={1},
  pages={1--10},
  year={2015},
  publisher={Nature Publishing Group}
}

@article{cooper2007behavioural,
  title={The behavioural inhibition system and behavioural approach system (BIS/BAS) scales: Measurement and structural invariance across adults and adolescents},
  author={Cooper, Andrew and Gomez, Rapson and Aucote, Helen},
  journal={Personality and individual differences},
  volume={43},
  number={2},
  pages={295--305},
  year={2007},
  publisher={Elsevier}
}

@article{leventhal2015measuring,
  title={Measuring anhedonia in adolescents: a psychometric analysis},
  author={Leventhal, Adam M and Unger, Jennifer B and Audrain-McGovern, Janet and Sussman, Steve and Volk, Heather E and Strong, David R},
  journal={Journal of personality assessment},
  volume={97},
  number={5},
  pages={506--514},
  year={2015},
  publisher={Taylor \& Francis}
}

@article{merikangas2007lifetime,
  title={Lifetime and 12-month prevalence of bipolar spectrum disorder in the National Comorbidity Survey replication},
  author={Merikangas, Kathleen R and Akiskal, Hagop S and Angst, Jules and Greenberg, Paul E and Hirschfeld, Robert MA and Petukhova, Maria and Kessler, Ronald C},
  journal={Archives of general psychiatry},
  volume={64},
  number={5},
  pages={543--552},
  year={2007},
  publisher={American Medical Association}
}

@article{merikangas2012true,
  title={The ‘true’ prevalence of bipolar II disorder},
  author={Merikangas, Kathleen R and Lamers, Femke},
  journal={Current opinion in psychiatry},
  volume={25},
  number={1},
  pages={19--23},
  year={2012},
  publisher={LWW}
}

@article{kessler2012twelve,
  title={Twelve-month and lifetime prevalence and lifetime morbid risk of anxiety and mood disorders in the United States},
  author={Kessler, Ronald C and Petukhova, Maria and Sampson, Nancy A and Zaslavsky, Alan M and Wittchen, Hans-Ullrich},
  journal={International journal of methods in psychiatric research},
  volume={21},
  number={3},
  pages={169--184},
  year={2012},
  publisher={Wiley Online Library}
}

@article{yarrington2021impact,
  title={Impact of the COVID-19 Pandemic on Mental Health among 157,213 Americans},
  author={Yarrington, Julia S and Lasser, Jana and Garcia, David and Vargas, Jose Hamilton and Couto, Diego Dotta and Marafon, Thiago and Craske, Michelle G and Niles, Andrea N},
  journal={Journal of Affective Disorders},
  year={2021},
  publisher={Elsevier}
}

@article{harle2017anhedonia,
  title={Anhedonia and anxiety underlying depressive symptomatology have distinct effects on reward-based decision-making},
  author={Harl{\'e}, Katia M and Guo, Dalin and Zhang, Shunan and Paulus, Martin P and Yu, Angela J},
  journal={PloS one},
  volume={12},
  number={10},
  pages={e0186473},
  year={2017},
  publisher={Public Library of Science San Francisco, CA USA}
}

@article{garrett2018updating,
  title={Updating beliefs under perceived threat},
  author={Garrett, Neil and Gonz{\'a}lez-Garz{\'o}n, Ana Mar{\'\i}a and Foulkes, Lucy and Levita, Liat and Sharot, Tali},
  journal={Journal of Neuroscience},
  volume={38},
  number={36},
  pages={7901--7911},
  year={2018},
  publisher={Soc Neuroscience}
}

@article{daw2011model,
  title={Model-based influences on humans' choices and striatal prediction errors},
  author={Daw, Nathaniel D and Gershman, Samuel J and Seymour, Ben and Dayan, Peter and Dolan, Raymond J},
  journal={Neuron},
  volume={69},
  number={6},
  pages={1204--1215},
  year={2011},
  publisher={Elsevier}
}

@article{king2018random,
  title={Random responses inflate statistical estimates in heavily skewed addictions data},
  author={King, Kevin M and Kim, Dale S and McCabe, Connor J},
  journal={Drug and alcohol dependence},
  volume={183},
  pages={102--110},
  year={2018},
  publisher={Elsevier}
}
\end{filecontents}
\addbibresource{main.bib}

% ----------------------------------%
% Define author / abstract block
% ----------------------------------%

@@

% ----------------------------------%
% Main text
% ----------------------------------%

\begin{document}

@@

% make abstract
\abstract{
\noindent Although online samples have many advantages for psychiatric research, some potential pitfalls of this approach are not widely understood. Here, we detail circumstances in which spurious correlations may arise between task behavior and symptom scores. The problem arises because many psychiatric symptom surveys have asymmetric score distributions in the general population, meaning that careless responders on these surveys will show apparently elevated symptom levels. If these participants are similarly careless in their task performance, this may result in a spurious association between symptom scores and task behavior. We demonstrate this pattern of results in two samples of participants (total N = 779) recruited online who performed one of two common cognitive tasks. False-positive rates for these spurious correlations increase with sample size, contrary to common assumptions. Excluding participants flagged for careless responding on surveys abolished these spurious correlations, but exclusion based on task performance alone was less effective.
}

@@

\section*{Introduction}

@@

% \subsection*{Screening for C/IE responding is common for task behavior but not for self-report surveys}
\subsection*{Narrative review of task and self-report screening practices}

@@

Among studies that we reviewed, approximately 80\% (39/49) used at least one method to identify C/IE responding in task behavior (see Table \ref{tab:tab01}). Of these, just over half relied on a single screening method, with considerable heterogeneity in behavior screening methods across studies. Most common (46\% of all studies) was identifying participants whose performance was statistically indistinguishable from chance-level on some measure of accuracy. Almost as common (38\%) was screening based on low response variability (i.e., excluding participants who predominantly responded in the same fashion across trials, such as using only a single response key).

@@

% \subsection*{Careless participants appear symptomatic when the overall level of symptom endorsement is low}
\subsection*{C/IE participants appear psychiatric when symptoms are rare}

@@

We hypothesise that spurious behavior-symptom correlations may emerge due to a mean-shift in the average level of symptom endorsement in participants engaging in C/IE responding relative to attentive participants. In turn, a mean-shift is expected to occur when the overall rate of symptom endorsement is low; that is, comparably higher scores are more likely for C/IE participants responding at random on a right skewed questionnaire. In line with our predictions, the average level of symptom endorsement was noticeably exaggerated in C/IE-responding participants for the symptom measures where symptom scores were most positively-skewed (7-up, 7-down, GAD-7; see Figure \ref{fig:fig01}). In contrast, where there was higher rates of symptom endorsement overall, the distributions of symptom scores between the two groups of participants were less noticeably distinct. Permutation testing confirmed that observed mean-shifts in symptom scores for C/IE participants were statistically significant for the majority of symptom measures (Table \ref{tab:tab02}).

Hereafter, we use the infrequency-item method as a primary means of identifying C/IE responding in our data. To verify this approach, we conducted three validation analyses. The first analysis compared estimated internal consistency of self-report measures between the C/IE and attentive groups. The logic is that, if C/IE responding manifests as a tendency to respond randomly, we should expect to see a decrease in the consistency of a measure in the C/IE responding group \cite{maniaci2014caring, desimone2018dirty, desimone2018differential}. In line with this reasoning, we observed a reduction in Cronbach's $\alpha$ in the C/IE group for the majority of survey instruments (Table \ref{tab:tab02}). A permutation test confirmed that the average decrease in internal consistency across measures was greater than would be expected by chance given the difference in participant numbers between groups (two-tailed, paired-samples $t$-test: $t(6)=3.689$, $p=0.021$, $d=1.506$, 95\% CI = [0.048, 0.141]).

@@

% \subsection*{Spurious symptom-behavior correlations produced by C/IE responding}
\subsection*{C/IE responding yields spurious symptom-behavior correlations}

@@

% \subsection*{Individuals with and without diagnosed psychiatric disorders fail attention checks at similar rates}
\subsection*{Psychiatric patients do not fail attention checks more often}
